A Detailed Analysis of 2020 U.S. Presidential Election: A Demographic Analysis Through Multilevel Regression and Post-Stratification

### 2024-03-16 paper/paper.qmd
minor changes

```{r}
#| echo: false
#| message: false
#| warning: false

@@

```{r} 
#| echo: false
#| message: false
#| warning: false
#| label: fig-distribution-by-state
#| fig-cap: Distribution of Sample and Population by Education

@@

# a state plot of the result
stateprops_plot %>% ggplot(aes(y = reorder(state, trump_predict), x = trump_predict, 
                               color = support)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
  labs(title = "Estimated Proportion of State Voting for Donal Trump",
       x = "Percentage", y= "State",
       color = "Support Trump") +
  scale_x_continuous(labels = scales::percent) +
  theme(axis.text.y = element_text(size = 5)) +
  scale_color_manual(values = c("Blue", "Red"))
```

@fig-distribution-by-state shows us the estimated of proportion of support for Trump or not by state using MRP with the inclusion of error terms. Each dot represents the point estimate of the proportion of support for Trump (red) or support for any other candidates including Joe Biden (blue) in each state. Horizontal lines extending from the dots represent error bars for these estimates. The length of each line indicates the uncertainty associated with each estimate. For instance, we can see that this uncertainty lies between 55% to slightly higher than 80% for Trump in MT (Massachusetts). The dashed green line in the middle at the 50% mark represents the threshold for majority support. On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest to the lowest.

@@

# Aggregating data by education level
post_props_education <- post_strat_data %>% 
  group_by(education, race, age_group, gender, state) %>%  # Adjust based on variables of interest
  summarise(n = n(), .groups ='drop') %>%
  group_by(education) %>%
  mutate(prop = n/sum(n))

# Generate predictions for each grouped cell based on education
post_props_education$estimate <- predict(survey_model, newdata = post_props_education, type = "response")

# Calculate the number of voters from each cell voting for Trump
post_props_education <- post_props_education %>% mutate(num_voters = n * estimate)

# Get errors for the prediction
errors_education <- predict(survey_model, newdata = post_props_education,
                                type = "response", se.fit = TRUE)
lower_education <- errors_education$fit - errors_education$se.fit
upper_education <- errors_education$fit + errors_education$se.fit

# Combine the proportions and the errors generated
post_props_education_error <- cbind(post_props_education, lower_education, upper_education) %>% rename(lower = ...10, upper = ...11)

# Aggregate predictions by education level
vote_trump_education <- post_props_education_error %>%
  mutate(trump_prop = estimate * prop,
         trump_prop_lower = lower * prop,
         trump_prop_upper = upper * prop) %>%
  group_by(education) %>%
  summarise(trump_predict = sum(trump_prop),
            trump_prop_lower = sum(trump_prop_lower),
            trump_prop_upper = sum(trump_prop_upper))

vote_trump_education %>% ggplot(aes(y = reorder(education, trump_predict), x = trump_predict)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
  labs(title = "Estimated Proportion of Voters for Trump by Education Level",
       x = "Percentage", y= "Education",
       color = "Support") +
  scale_x_continuous(labels = scales::percent) +
  theme(axis.text.y = element_text(size = 5)) +
  scale_color_manual(values = c("Blue", "Red"))

>>>>

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-distribution-by-state
#| fig-cap: Distribution of Sample and Population by State

@@

@@

# define colors 
support_colors <- c("No" = "Blue", "Yes" = "Red")

# state plot
ggplot(stateprops_plot, aes(y = reorder(state, trump_predict), 
                            x = trump_predict, color = support)) +
  geom_point() +
  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "orange") +
  scale_x_continuous(labels = scales::percent) +
  labs(title = "Estimated Proportion of State Voting for Donald Trump",
       x = "Percentage", y= "State",
       color = "Candidate") +
  theme(axis.text.y = element_text(size = 5), legend.position = "right") +
  scale_color_manual(values = c("Blue", "Red"))
```

@fig-distribution-by-state shows us the estimated of proportion of support for Trump or not by state using MRP with the inclusion of error terms. Each dot represents the point estimate of the proportion of support for Trump (red) or support for any other candidates including Joe Biden (blue) in each state. Horizontal lines extending from the dots represent error bars for these estimates. The length of each line indicates the uncertainty associated with each estimate. For instance, we can see that this uncertainty lies between 55% to slightly higher than 80% for Trump in MT (Massachusetts). The dashed orange line in the middle at the 50% mark represents the threshold for majority support. On the y-axis, each state is listed with its abbreviations and is ordered based on the proportion of support for Trump from the highest to the lowest.

@@

# group by education level
post_props_education <- post_strat_data %>% 
  group_by(education, race, age_group, gender, state) %>%  
  summarise(n = n(), .groups ='drop') %>%
  group_by(education) %>%
  mutate(prop = n/sum(n))

# predictions based on education
post_props_education$estimate <- predict(survey_model, newdata = post_props_education, type = "response")

# number of voters for Trump
post_props_education <- post_props_education %>% mutate(num_voters = n * estimate)

# errors for the prediction
errors_education <- predict(survey_model, newdata = post_props_education,
                                type = "response", se.fit = TRUE)
lower_education <- errors_education$fit - errors_education$se.fit
upper_education <- errors_education$fit + errors_education$se.fit

# combine the proportions and the errors
post_props_education_error <- cbind(post_props_education, lower_education, upper_education) %>% 
  rename(lower = ...10, upper = ...11)

# predictions by education level
vote_trump_education <- post_props_education_error %>%
  mutate(trump_prop = estimate * prop,
         trump_prop_lower = lower * prop,
         trump_prop_upper = upper * prop) %>%
  group_by(education) %>%
  summarise(trump_predict = sum(trump_prop),
            trump_prop_lower = sum(trump_prop_lower),
            trump_prop_upper = sum(trump_prop_upper))

# education plot
  ggplot(vote_trump_education, aes(y = reorder(education, trump_predict), 
                                   x = trump_predict)) +
  geom_point(stat = "identity") +
  geom_errorbar(aes(xmin = trump_prop_lower, xmax = trump_prop_upper, width = 0.2)) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "green") +
  scale_x_continuous(labels = scales::percent) +
  labs(title = "Estimated Proportion of Voters for Trump by Education",
       x = "Percentage", y= "Education") + 
  theme_minimal() + theme(axis.text.y = element_text(size = 12))
